The reference model approach in feature selection problems \vspace{.1in}

### 2019-07-17 paper/ref_approach.tex
section 2: included more references

% In Figure \ref{fig:correlation} the reference model filters the noise.
% This simple idea is what happens every time a reference model is used, as it is evident for example in the preconditioning by \cite{paul2008preconditioning} or in the draw-by-draw projection for GLMs by \cite{paper:original_proj}.\\
% Above the correlation was used to estimate the relevance of a covariate.  
As an another example, let us consider a candidate submodel in a feature selection process and let us refer to the parameter distribution of the submodel by $\pi$ and to the induced predictive distribution by $q_{\pi}(\tilde{y})$. We would like to choose $\pi$ so that $q_{\pi}(\tilde{y})$ maximises some predictive performance utility, for example, the expected log-predictive density (elpd) defined as:
\
\begin{equation}\label{eq:elpd}
\text{elpd}[q_{\pi}]=\int \text{log}\,q_{\pi}(\tilde{y})p_{t}(\tilde{y})d\tilde{y},
\end{equation}
where $p_{t}(\tilde{y})$ denotes the (usually unknown) true generating mechanism of future data $\tilde{y}$. If we refer to the posterior predictive distribution of a reference model with $p(\tilde{y}|D)$, where $D$ stands for the data on which we conditioned on, we can approximate \eqref{eq:elpd} using $p(\tilde{y}|D)$ instead of the true data generation mechanism $p_{t}(\tilde{y})$. The maximisation of the elpd using the reference model predictive distribution is equivalent to the minimisation of the KL divergence between the reference model's and the submodel's predictive distributions:
\
\begin{equation} \label{proj_as_filter}
\underset{\pi}{\text{max}} \; \int \text{log}\,q_{\pi}(\tilde{y})p(\tilde{y}|D)d\tilde{y} \leftrightarrow \underset{\pi}{\text{min}} \; \text{KL}[p(\tilde{y}|D)||q_{\pi}(\tilde{y})] 
\end{equation}
The term on the right-hand side of Equation \eqref{proj_as_filter} describes what is referred to as the projection of the predictive distribution, which is the general idea behind the projection predictive approach (cite). Again the reference model is acting as a filter, or substitute, on data.

@@

Relying on the idea of the reference model as a noise-filter, a generic reference model approach, which can be used on top of any feature selection procedure, consists of simply substituting the target observations with corresponding point predictions of the reference model, for instance, the posterior predictive means. Afterwards, any method for feature selection can be used. Note that we do not argue that such a reference model approach should necessarily be used as is for feature selection in real applications. There are several more sophisticated reference model approaches available in the literature (cite), which are likely to yield even better results. Rather, the reason for focussing on this simple version is to enable a fair comparison between a reference model and a purely data based approach when otherwise using the same selection procedures. Other kinds of comparisons, as the one presented in Section \ref{projection}, work well as motivational examples but fail in terms of generality since the presence or absence of a reference model is not the only aspect differentiating the two procedures. We will come back to the more general comparison in Section \ref{comparison}.

@@

As a motivational example to the implications of using a reference model in feature selection problems, we show its application in the form of the projection predictive approach, which uses the reference model to project the posterior samples to the submodels (cite). The \texttt{R}-package \texttt{projpred} (available at \url{https://CRAN.R-project.org/package=projpred}) implements the projection in case of submodels in the family of the generalised linear models and, additionally, it provides a framework to choose the optimal submodel size through a predictive utility comparison between the submodels and the reference model. 
\\
We now summarise the workflow of the projection predictive approach in the particular case of the draw-by-draw projection \cite[original formulation by][]{paper:original_proj}, following \cite{paper:projpred}. Suppose we have observed $n$ statistical units with target values $\{y_{i}\}_{i=1}^{n}$ and a set of observed covariates for which we want to obtain a minimally relevant subset. Than, the main steps are the following:
\begin{enumerate}
\item Devise and fit a reference model. Let $\{\boldsymbol{\theta}_{*}^{s}\}_{s=1}^{S}$ be the set of $S$ posterior samples of the reference model's parameters.
\item Rank the covariates according to their relevance using some heuristics and consider as candidate submodels only those which preserve this order,  starting from including only the highest ranked covariate (the submodels are then naturally identified by their model size). This is not strictly necessary but reduces the number of submodels to consider in the following steps.
\item For each submodel, project the reference model's samples as follows:
\begin{equation}\label{eq:draw_by_draw}
\boldsymbol{\theta}_{\perp}^{s} = \underset{\boldsymbol{\theta^{s}}\in\Theta}{\text{argmin}} \frac{1}{n}\sum_{i=1}^{n}\text{KL} \left( p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})\;||\;q(\tilde{y}_{i}|\boldsymbol{\theta}^{s}) \right) \quad s=1,..,S
\end{equation}
where $p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})$ stands for the predictive distribution of the reference model with parameters fixed at the posterior sample $\boldsymbol{\theta}_{*}^{s}$ and conditioning on all the covariate values related to the statistical unit (identified by the subscript $i$), whereas $q(\tilde{y}_{i}|\boldsymbol{\theta}^{s})$ is the predictive distribution of the submodel. The projected samples $\boldsymbol{\theta}_{\perp}^{s}$ then correspond to the submodel's parameter distribution.
\item For each submodel (size), test the predictive performance for instance via cross-valdiation.
\item Choose the smallest submodel (size) that is sufficiently close to the reference model's predictive utility score.
\end{enumerate}

Expression \eqref{eq:draw_by_draw} is not an easy optimisation problem in general, but in the special case of submodels for the family of generalize lineder models, \eqref{eq:draw_by_draw} reduces to a maximum likelihood estimation problem, which can be easily optimised (cite). For further details on the projective prediction workflow and implementation see \cite{paper:projpred}.

@@

\cite{paper:bodyfat} report results using steplm with a significance level of 0.157 with AIC selection, fixing abdomen and height to be always included in the model. Here, we implement steplm via AIC without fixing any variable. For selection via projpred, the reference model includes all the covariates, we regularize model complexity with a regularised horseshoe prior on the covariate coefficients (cite), submodels are explored using forward search, and the predictive utility is the expected log-predictive density (elpd) estimated using PSIS-LOO \citep{paper:psis_loo}. The experiments are implemented in \texttt{R} (cite). The steplm approach is carried out combining the \texttt{step} and \texttt{lm} functions, whereas the selection via projpred is done with the \texttt{projpred} package. In order to automatise the latter, we select the submodel as the one with the smallest size which has an elpd score higher (i.e., better) than the reference model score with probability of at least 5\%. The inclusion frequencies of each feature in the final model after 100 bootstrap samples are depicted in Figure \ref{fig:inclusion_frequencies}. The projpred approach has only two variables, 'abdomen' and 'weight', with inclusion frequencies above 50\% ('abdomen' is the only included always), the third most included is `wrist' at 44\%, and the fourth is `height' at 35\%, while steplm has seven features above 50\%. Such a lower stability of steplm can be observed also in the bootstrap model selection frequencies reported in Table \ref{tab:model_frequencies}. For exaple, the first five selected models have a cumulative frequency of 76\% with projpred, but only of 14\% with steplm. In addition, we note that the size of the selected model is much smaller when using projpred.

@@

Table \ref{tab:model_performances} reports the predictive performances, in terms of root mean square error, of the full model and the corresponding selected submodel using both projpred and steplm. We also repeat the selection after adding truely uncorrelated (i.e., irrelevant) features up to a total of 100 covariates. For this experiment, we report both the size of the selected submodel and the number of selected irrelevant features. Whilst, with the original data, we do not observe any gain using projpred in terms of predictive performance, what we do observe is a smaller size for the selected submodel, at the cost of a higher computation time compared to steplm. However, when additional irrelevant features are added, projpred is able to notably improve the predictive performance keeping the same submodel size, whereas steplm selects a much larger submodel also including a large number of irrelvant features (13 irrelevant out of a total of 19 selected features).

@@

We consider the reference model approach as a family of methods, thus providing a general answer to the question of computational burden is not possible because not unique. It is possible to identify a common and inevitable overhead, though, which is the cost of fitting the reference model. In the simple ''bare-bones'' reference model approach we have described above, it is actually the only added cost. In general, a full Bayesian model is advisable in order to have the best predictive performance. However, including all the available features can often be quite computationally demanding. In such a case, a first possible speed-up could be using screening or dimensionality reduction techniques, as for example the supervised principal components \citep{bair2006prediction,piironen2018}. However, in some cases, this does not always result in an appreciable speed-up. For example, when there are several predictive but mutually uncorrelated features, all of them  have to be included in the reference model anyway. Other alternatives could be fixing some parameters to their maximum marginal likelihood value (typically referred as empirical Bayes; cite) or use a completely frequentist approach, if predictions are good enough.

>>>>

% In Figure \ref{fig:correlation} the reference model filters the noise.
% This simple idea is what happens every time a reference model is used, as it is evident for example in the preconditioning by \cite{paul2008preconditioning} or in the draw-by-draw projection for GLMs by \cite{paper:original_proj}.\\
% Above the correlation was used to estimate the relevance of a covariate.  
As an another example, let us consider a candidate submodel in a feature selection process and let us refer to the parameter distribution of the submodel by $\pi$ and to the induced predictive distribution by $q_{\pi}(\tilde{y})$. We would like to choose $\pi$ so that $q_{\pi}(\tilde{y})$ maximises some predictive performance utility, for example, the expected log-predictive density (elpd) defined as:
\
\begin{equation}\label{eq:elpd}
\text{elpd}[q_{\pi}]=\int \text{log}\,q_{\pi}(\tilde{y})p_{t}(\tilde{y})d\tilde{y},
\end{equation}
where $p_{t}(\tilde{y})$ denotes the (usually unknown) true generating mechanism of future data $\tilde{y}$. If we refer to the posterior predictive distribution of a reference model with $p(\tilde{y}|D)$, where $D$ stands for the data on which we conditioned on, we can approximate \eqref{eq:elpd} using $p(\tilde{y}|D)$ instead of the true data generation mechanism $p_{t}(\tilde{y})$. The maximisation of the elpd using the reference model predictive distribution is equivalent to the minimisation of the KL divergence between the reference model's and the submodel's predictive distributions:
\
\begin{equation} \label{proj_as_filter}
\underset{\pi}{\text{max}} \; \int \text{log}\,q_{\pi}(\tilde{y})p(\tilde{y}|D)d\tilde{y} \leftrightarrow \underset{\pi}{\text{min}} \; \text{KL}[p(\tilde{y}|D)||q_{\pi}(\tilde{y})] 
\end{equation}
The term on the right-hand side of Equation \eqref{proj_as_filter} describes what is referred to as the projection of the predictive distribution, which is the general idea behind the projection predictive approach \cite[see][]{paper:projpred}. Again the reference model is acting as a filter, or substitute, on data.

@@

Relying on the idea of the reference model as a noise-filter, a generic reference model approach, which can be used on top of any feature selection procedure, consists of simply substituting the target observations with corresponding point predictions of the reference model, for instance, the posterior predictive means. Afterwards, any method for feature selection can be used. Note that we do not argue that such a reference model approach should necessarily be used as is for feature selection in real applications. There are several more sophisticated reference model approaches available in the literature \cite[see, e.g.][]{paper:projpred,paananen2017variable,paul2008preconditioning}, which are likely to yield even better results. Rather, the reason for focussing on this simple version is to enable a fair comparison between a reference model and a purely data based approach when otherwise using the same selection procedures. Other kinds of comparisons, as the one presented in Section \ref{projection}, work well as motivational examples but fail in terms of generality since the presence or absence of a reference model is not the only aspect differentiating the two procedures. We will come back to the more general comparison in Section \ref{comparison}.

@@

As a motivational example to the implications of using a reference model in feature selection problems, we show its application in the form of the projection predictive approach, which uses the reference model to project the posterior samples to the submodels \citep{paper:original_proj}. The \texttt{R}-package \texttt{projpred} (available at \url{https://CRAN.R-project.org/package=projpred}) implements the projection in case of submodels in the family of the generalised linear models and, additionally, it provides a framework to choose the optimal submodel size through a predictive utility comparison between the submodels and the reference model. 
\\
We now summarise the workflow of the projection predictive approach in the particular case of the draw-by-draw projection \cite[original formulation by][]{paper:original_proj}, following \cite{paper:projpred}. Suppose we have observed $n$ statistical units with target values $\{y_{i}\}_{i=1}^{n}$ and a set of observed covariates for which we want to obtain a minimally relevant subset. Than, the main steps are the following:
\begin{enumerate}
\item Devise and fit a reference model. Let $\{\boldsymbol{\theta}_{*}^{s}\}_{s=1}^{S}$ be the set of $S$ posterior samples of the reference model's parameters.
\item Rank the covariates according to their relevance using some heuristics and consider as candidate submodels only those which preserve this order,  starting from including only the highest ranked covariate (the submodels are then naturally identified by their model size). This is not strictly necessary but reduces the number of submodels to consider in the following steps.
\item For each submodel, project the reference model's samples as follows:
\begin{equation}\label{eq:draw_by_draw}
\boldsymbol{\theta}_{\perp}^{s} = \underset{\boldsymbol{\theta^{s}}\in\Theta}{\text{argmin}} \frac{1}{n}\sum_{i=1}^{n}\text{KL} \left( p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})\;||\;q(\tilde{y}_{i}|\boldsymbol{\theta}^{s}) \right) \quad s=1,..,S
\end{equation}
where $p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})$ stands for the predictive distribution of the reference model with parameters fixed at the posterior sample $\boldsymbol{\theta}_{*}^{s}$ and conditioning on all the covariate values related to the statistical unit (identified by the subscript $i$), whereas $q(\tilde{y}_{i}|\boldsymbol{\theta}^{s})$ is the predictive distribution of the submodel. The projected samples $\boldsymbol{\theta}_{\perp}^{s}$ then correspond to the submodel's parameter distribution.
\item For each submodel (size), test the predictive performance for instance via cross-valdiation.
\item Choose the smallest submodel (size) that is sufficiently close to the reference model's predictive utility score.
\end{enumerate}

Expression \eqref{eq:draw_by_draw} is not an easy optimisation problem in general, but in the special case of submodels for the family of generalised linear models, \eqref{eq:draw_by_draw} reduces to a maximum likelihood estimation problem, which can be easily optimised \citep{paper:original_proj}. For further details on the projective prediction workflow and implementation see \cite{paper:projpred}.

@@

\cite{paper:bodyfat} report results using steplm with a significance level of 0.157 with AIC selection \citep{akaike1974new}, fixing abdomen and height to be always included in the model. Here, we implement steplm via AIC without fixing any variable. For selection via projpred, the reference model includes all the covariates, we regularize model complexity with a regularised horseshoe prior on the covariate coefficients \citep{paper:rhs}, submodels are explored using forward search, and the predictive utility is the expected log-predictive density (elpd) estimated using PSIS-LOO \citep{paper:psis_loo}. The experiments are implemented in \texttt{R} \citep{Rcore2018}. The steplm approach is carried out combining the \texttt{step} and \texttt{lm} functions, whereas the selection via projpred is done with the \texttt{projpred} package. In order to automatise the latter, we select the submodel as the one with the smallest size which has an elpd score higher (i.e., better) than the reference model score with probability of at least 5\%. The inclusion frequencies of each feature in the final model after 100 bootstrap samples are depicted in Figure \ref{fig:inclusion_frequencies}. The projpred approach has only two variables, 'abdomen' and 'weight', with inclusion frequencies above 50\% ('abdomen' is the only included always), the third most included is `wrist' at 44\%, and the fourth is `height' at 35\%, while steplm has seven features above 50\%. Such a lower stability of steplm can be observed also in the bootstrap model selection frequencies reported in Table \ref{tab:model_frequencies}. For exaple, the first five selected models have a cumulative frequency of 76\% with projpred, but only of 14\% with steplm. In addition, we note that the size of the selected model is much smaller when using projpred.

@@

Table \ref{tab:model_performances} reports the predictive performances, in terms of root mean square error, of the full model and the corresponding selected submodel using both projpred and steplm. We also repeat the selection after adding truly uncorrelated (i.e., irrelevant) features up to a total of 100 covariates. For this experiment, we report both the size of the selected submodel and the number of selected irrelevant features. Whilst, with the original data, we do not observe any gain using projpred in terms of predictive performance, what we do observe is a smaller size for the selected submodel, at the cost of a higher computation time compared to steplm. However, when additional irrelevant features are added, projpred is able to notably improve the predictive performance keeping the same submodel size, whereas steplm selects a much larger submodel also including a large number of irrelvant features (13 irrelevant out of a total of 19 selected features).

@@

We consider the reference model approach as a family of methods, thus providing a general answer to the question of computational burden is not possible because not unique. It is possible to identify a common and inevitable overhead, though, which is the cost of fitting the reference model. In the simple ''bare-bones'' reference model approach we have described above, it is actually the only added cost. In general, a full Bayesian model is advisable in order to have the best predictive performance. However, including all the available features can often be quite computationally demanding. In such a case, a first possible speed-up could be using screening or dimensionality reduction techniques, as for example the supervised principal components \citep{bair2006prediction,piironen2018}. However, in some cases, this does not always result in an appreciable speed-up. For example, when there are several predictive but mutually uncorrelated features, all of them  have to be included in the reference model anyway. Other alternatives could be fixing some parameters to their maximum marginal likelihood value \cite[typically referred as empirical Bayes;][]{casella1985introduction} or use a completely frequentist approach, if predictions are good enough.
